---
title: "Dave's Evenness Paper"
author: "Diego Rincon & Riley Anderson"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
graphics: yes
output:
  github_document:
    toc: yes
    toc_depth: 5
    pandoc_args: --webtex
  html_document:
    keep_md: yes
    theme: readable
    mathjax: default
  html_notebook:
    code_folding: hide
    theme: readable
    mathjax: default
  pdf_document:
    toc: yes
header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
editor_options:
  chunk_output_type: console
---

```{r setup, include = F}
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Knitr Options
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Set root directory to the project directory
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())


# Set default knitr options: 
# Suppress warnings and messages, cache chunks, 
#  set default figure size to 6x8 at 300 dpi, and save a png and pdf
knitr::opts_chunk$set(warning = F, message = F, collapse = T, cache = T,
    fig.height = 6, fig.width = 8, dpi = 300, # 6x8" @ 300dpi:1800x2400=4.3MP
    dev = c('png', 'pdf'), dev.args = list(pdf = list(onefile = F)))

```



## Overview

What is this analysis about?

### Summary of Results
* 

```{r Main_Code, include = F, cache = F}

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Setup - This code is run, but output is hidden
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Load Packages
library(tidyverse) # Needed for data wrangling: dplyr, tidyr, ggplot2
library(cowplot) # Needed for publication-quality ggplots
library(nls.multstart)
require(bbmle)
library(chemodiv)

# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@
# @@@@@ Data Preparation
# @@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@

# Import datasets
dave <- read.csv("data/dave.csv")
block1 <- read.csv("data/block1.csv")
block2 <- read.csv("data/block2.csv")
block3 <- read.csv("data/block3.csv")

```

```{r data_wrangling, echo = F}

#Dataset with only perfect evenness treatments
alleven <- rbind(block1[136:165, ], block2[136:165, ], block3[91:110, ])

alldata <- rbind(block1, block2, block3)

alldata$Even <- calcDiv(alldata[,2:4], type = "HillEven", q = 1)

colnames(alldata)[7] <- "Even"

```

# Diego's model with binomial pdf

This model follows Northfield's equation:

$$Y = m_{b} (p_{0} + p_{1}e^{-a_{1}x_{1}} + p_{2}e^{-a_{2}x_{2}}+p_{3}e^{-a_{3}x_{3}})$$

```{r Diego_code, echo = F, comment = ""}



# Likelihood function

LL1 <- function(mb, p0, p1, p2, p3, a1, a2, a3) {
  Y = mb * (p0 + (p1 * exp(-a1 * x1)) + (p2 * exp(-a2 * x2)) + (p3 * exp(-a3 * x3))) # this is Tobin's function
  -sum(dbinom(resp, size = initial, prob = Y, log = TRUE)) # Decided to use a binomial
}

# Optimization

# Makes more sense because of the boundaries in estimation, but does not work

# MLL1 <- mle2(LL1, start = list(mb = 0.5, p0 = 0.1, p1 = 0.3, p2 = 0.3, p3 = 0.3, a1 = 0.001, a2 = 0.001, a3 = 0.001), 
#              data = list(x1 = alleven$Sc, x2 = alleven$Hm, x3 = alleven$Bb, resp = alleven$Alive, initial = rep(10, 80)),
#              lower = list(p0 = 0, p1 = 0, p2 = 0, p3 = 0), 
#              upper = list(p0 = 1, p1 = 1, p2 = 1, p3 = 1), method = "L-BFGS-B")

# No boundaries but works (kind of..)
MLL1a <- mle2(LL1, start = list(mb = 0.5, p0 = 0.1, p1 = 0.3, p2 = 0.3, p3 = 0.3, a1 = 0.001, a2 = 0.001, a3 = 0.001), 
              data = list(x1 = alleven$Sc, x2 = alleven$Hm, x3 = alleven$Bb, resp = alleven$Alive, initial = rep(10, 80)),
              method = "Nelder-Mead")

summary(MLL1a)

```


# Diego's model with poisson pdf

This works with the boundaries in estimation
```{r poisson_MLE, echo = F}



# Likelihood function

LL2 <- function(mb, p0, p1, p2, p3, a1, a2, a3) {
  Y = mb * (p0 + (p1 * exp(-a1 * x1)) + (p2 * exp(-a2 * x2)) + (p3 * exp(-a3 * x3)))
  -sum(dpois(resp, lambda = Y, log = TRUE)) # trying a poisson dist.
}

# Optimization

MLLp1 <- mle2(LL2, start = list(mb = .5, p0 = 0, p1 = 1, p2 = 1, p3 = 1, a1 = 0.001, a2 = 0.001, a3 = 0.001), 
             data = list(x1 = alleven$Sc, x2 = alleven$Hm, x3 = alleven$Bb, resp = alleven$Alive, initial = rep(10, 80)),
             lower = list(p0 = 0, p1 = 0, p2 = 0, p3 = 0), 
             upper = list(p0 = 10, p1 = 10, p2 = 10, p3 = 10), method = "L-BFGS-B")

summary(MLLp1)


est_params <- data.frame(as.list(coef(MLLp1)))


# Create a data frame for prediction
prediction_df <- data.frame(
  x1 = alleven$Sc,
  x2 = alleven$Hm,
  x3 = alleven$Bb
)

# Add columns for fitted values using the estimated parameters
prediction_df$fitted_values <- with(est_params, 
  mb * (p0 + (p1 * exp(-a1 * prediction_df$x1)) + (p2 * exp(-a2 * prediction_df$x2)) + (p3 * exp(-a3 * prediction_df$x3)))
)

# Make a figure
ggplot(prediction_df, aes(x = x1, y = fitted_values)) +
  geom_point() +
  labs(x = "Predator abundance",
       y = "Prey alive") +
  theme_cowplot() +
  geom_smooth(method = "lm",
              formula = y ~ exp(-x * est_params$a1))


```




# Riley's model

This is not likely to be useful, it was more of a learning exercise to understand the nls model features. The main problem is that this model assumes the exponential regression coefficient is constant across each species. 

```{r data_generation, echo = F}

# Generate example data
set.seed(123)
pred.abun <- seq(100, 10000, length.out = 100)
evenness <- sample(1:4, 100, replace = TRUE)
alive <- 10 * exp(-10 * pred.abun) + 5 * evenness + 
  rnorm(100, mean = 0, sd = 5)

# Create a data frame
data <- data.frame(pred.abun = pred.abun, evenness = evenness, alive = alive)

```

# Model
A negative exponential model with variable asymptote:

$$y(t)=ae^{-bt}+cx$$

```{r define_model, echo = F}

# Define the model function
exp_decay <- function(pred.abun, a, b, c, evenness) {
  a * exp(-b * pred.abun) + c * evenness
}

```


```{r models, echo = F, eval=FALSE}

# Fit the model to the data
fit <- nls_multstart(
  alive ~ exp_decay(pred.abun, a, b, c, evenness),
                     data = data,
                     iter = 100,
                     start = list(a = 10, b = 2, c = 5),
                     start_lower = c(0, 0, 0),
                     upper = c(100, Inf, 10))

summary(fit)


# Extract the coefficients
a <- coef(fit)[1]
b <- coef(fit)[2]
c <- coef(fit)[3]

# Create a data frame for plotting the fitted curve
fitted_data <- data.frame(pred.abun = pred.abun, evenness = evenness,
          fitted_values = exp_decay(pred.abun, a, b, c, evenness))

data <- data %>% 
  mutate(fitted_values = fitted_data$fitted_values,
         evenness = factor(evenness))

```

```{r figure_Riley_mod, echo = F, eval = F}

ggplot(data) +
  geom_point(aes(x = pred.abun,
                 y = alive,
                 fill = evenness), shape = 1) +
  geom_smooth(aes(x = pred.abun,
                  y = fitted_values,
                  fill = evenness),
              stat = "identity") +
  labs(x = "Predator abundance", y = "Prey alive", fill = "Evenness") +
  theme_cowplot()

```


```{r Graph_name, echo = F}

# Put each graph in a separate code block with a different name

```


## Session Information

```{r Session_Info, echo = F, comment = ""}

# Add session information to help with reproduceability
sessionInfo()


```


